SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-02 docs/questioning.tex
work on runtime tainting

Runtime tainting is based on data-flow analysis.
In practice, it enforces security policies by marking untrusted
(``tainted") data and tracing its flow through the program.
Since information flow in a system cannot be verified by examining
a single execution trace of the system, the results of taint
analysis will necessarily reflect approximate information regarding
the information flow characteristics of the system to which it is applied.

>>>>

A runtime tainting scheme marks untrusted
(``tainted") data (for instance, a variable set
by a field in a web form) and traces its flow through
the program. If the variable is used in an expression
that sets another variable, that variable is also
marked as untrusted and so on. If any of these variables
is used to execute a potentially risky command (``sink") the
scheme may act accordingly.

Runtime tainting is provided as a feature in some
programming languages, such as Perl and Ruby.
By enabling the feature, Perl would refuse to run code
vulnerable to an {\sc sql} injection attack
(consider a tainted variable being used in a query)
and exit with an error message.

There are different implementations of these approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}). However, this is a simplistic
approach that may lead to many false positives.
{\sc csse}~\cite{PB05} and the scheme by {\it Xu et al.}~\cite{XBS06}
apply further checks when they identify that tainted data
have reached a sink. In particular, {\sc csse} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and other. When tainted data reaches a sink {\sc csse} performs
syntactic checks it based on the aforementioned metadata
(locus {\sc z}).

% As the author of {\sc csse} state~\cite{PB05}, the prevention
% of {\sc xss} attack requires a more complex analysis.
